---
title: 'Criminalization of Girls in California: A Fact Sheet'
author: "Shaina Mackin | The Vera Institute"
date: "March 29, 2022"
output:
  pdf_document: default
  github_document: default
---

```{r setup, echo=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(plotly)
library(patchwork)



theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r import, echo = FALSE, message = FALSE, warning = FALSE}
adp_df = read_excel("./data/juvenile_adp_2002-2021.xlsx", col_types = "text") 
```
<br>

<center>

# ACROSS THE STATE
From 2012-2020, **1,122** girls were detained in non-secure facilities and **32,331** in secure facilities.                                    
**18.4%**	of detentions between 2012-2017 were in secure facilities.

# TOP INCARCERATING COUNTIES

```{r top_incarcerators, echo=FALSE}
County <- c('Los Angeles', 'San Bernadino', 'Ventura', 'Orange', 'Contra Costa', 'Santa Clara', 'Santa Barbara', 'Kern', 'Fresno', 'Mendocino') 
Percent <- c('12.5%', '10.99%', '7.12%', '6.54%', '5.12%', '4.62%', '4.23%', '3.97%', '3.5%', '3.3%')
Count <- c(577, 509, 330, 303, 237, 214, 196, 184, 162, 153)

top_incarcerators <- data.frame(County, Percent, Count)

Counties <- c('Glenn', 'Modoc', 'Mendocino', 'Sonoma', 'Imperial', 'San Francisco', 'Shasta', 'Ventura', 'Santa Clara', 'San Luis Obispo') 
Frequency <- c(6,4,5,2,4,3,5,2,2,NA)
Percent_Cases <- c('100%','100%','56.7%','49.2%','48.7%','48.4%','47.8%','43.5%','41.6%','41.1%')

most_likely <- data.frame(Counties, Percent_Cases, Frequency)

```
<br>
<center>
## Counties that detained the most girls in 2017: 
</center>

```{r, echo=FALSE}
knitr::kable(top_incarcerators, col.names = c('County', 'Percent of State Total', 'Number Detained'), valign = 't')
```

<br>
<center>
## Counties most likely to detain girls in 2017:
</center>

```{r, echo=FALSE}
knitr::kable(most_likely, col.names = c('County', 'Percent of Cases', 'Number of Times in Top 10'), valign = 't')
```
<center>

# AVERAGE DAILY POPULATIONS
## Statewide by County
California's Board of State and Community Corrections (BSCC) reports Average Daily Populations (ADP) of juvenile facilities by county. In 2021, LA County's annual sum of monthly ADP counts was over twice that of Kern's, the second highest.
<br>

```{r tidy, echo=FALSE, message = FALSE, warning = FALSE}
adp_girls = adp_df %>%
  janitor::clean_names() %>%
  mutate(
    adp_pre_disposition_female = replace(adp_pre_disposition_female, adp_pre_disposition_female == "D", 0),
    adp_pre_disposition_female = replace(adp_pre_disposition_female, adp_pre_disposition_female == "d", 0),
     adp_pre_disposition_female = replace(adp_pre_disposition_female, adp_pre_disposition_female == "u", 0),
     adp_post_disposition_female = replace(adp_post_disposition_female, adp_post_disposition_female == "D", 0),
     adp_post_disposition_female = replace(adp_post_disposition_female, adp_post_disposition_female == "d", 0),
    adp_post_disposition_female = replace(adp_post_disposition_female, adp_post_disposition_female == "u", 0),
  ) %>%
  select(county, year, month, adp_pre_disposition_female, adp_post_disposition_female) %>%
  mutate(
    adp_pre_disposition_female = as.numeric(adp_pre_disposition_female),
    adp_post_disposition_female = as.numeric(adp_post_disposition_female)) %>%
  mutate(
    total_female = adp_pre_disposition_female + adp_post_disposition_female) %>%
    group_by(county, year, month) %>%
  summarise(
    adp_pre_disposition_female = sum(adp_pre_disposition_female),
    adp_post_disposition_female = sum(adp_post_disposition_female),
    total_female = sum(total_female)) %>%
  mutate(
    year = as.numeric(unlist(year)),
    month = as.numeric(unlist(month)),
    day = "01"
    ) %>%
  mutate(
    month = formatC(as.numeric(month), width=2,format='f',digits = 0, flag='0') 
  ) %>%
  select(county, day, month, year, total_female) %>%
  unite('the_date', day:year, remove = FALSE) %>%
  mutate(
    the_date = str_replace(the_date, "_", "\\")
  ) %>%
   mutate(
    the_date = str_replace(the_date, "_", "\\")
  ) %>%
  mutate(
    the_date = lubridate::dmy(the_date)
  )

write.csv(adp_girls,"/Users/shainamackin/Desktop/Vera/California Criminalization/data/adp_girls.csv")
```

```{r annual_adp, fig.height=3.2, echo = FALSE, message = FALSE}
annual_adp = adp_girls %>%
  select(county, month, year, total_female) %>%
  group_by(county, year) %>%
  summarize(annual = sum(total_female))

annual_2021_plot = annual_adp %>%
  filter(year == 2021) %>%
  ggplot(aes(x = reorder(county, -annual), y = annual, fill = annual)) +
  geom_bar(position="dodge", stat="identity") +
  theme(legend.position=c(0.96, 0.56)) +
  labs(
    title = "Annual Sum of Monthly ADP by CA County, 2021",
    x = "County",
    y = "Annual ADP Sum") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))

annual_2021_plot
```

```{r months, include = FALSE, echo=FALSE, message = FALSE, warning= FALSE}
jan_2021 = adp_girls %>%
  filter(year == 2021 & month == '01')

feb_2021 = adp_girls %>%
  filter(year == 2021 & month == '02')

mar_2021 = adp_girls %>%
  filter(year == 2021 & month == '03')

plot_jan_2021 = jan_2021 %>%
  mutate(ADP = total_female) %>%
  ggplot(aes(x = reorder(county, -total_female), y = ADP, fill = county)) +
  geom_bar(position="dodge", stat="identity") +
  labs(
    title = "Average Daily Population of Girls in Californian Facilities: Jan 1, 2021",
    x = "County",
    y = "ADP") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))

plot_feb_2021 = feb_2021 %>%
  mutate(ADP = total_female) %>%
  ggplot(aes(x = reorder(county, -total_female), y = ADP, fill = county)) +
  geom_bar(position="dodge", stat="identity") +
  labs(
    title = "Average Daily Population of Girls in Californian Facilities: Feb 1, 2021",
    x = "County",
    y = "ADP") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))

plot_mar_2021 = mar_2021 %>%
  mutate(ADP = total_female) %>%
  ggplot(aes(x = reorder(county, -total_female), y = ADP, fill = county)) +
  geom_bar(position="dodge", stat="identity") +
  labs(
    title = "Average Daily Population of Girls in Californian Facilities: Mar 1, 2021",
    x = "County",
    y = "ADP") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))
  
```
<br>

## Zooming in on Santa Clara County 
In 2020, Santa Clara County's highest ADP (n=**21**) occurred in January, with a steady decline continuing throughout the COVID-19 pandemic. By 2021, even the county's highest ADP, occurring in February, had seen an **81%** decrease (n=**3**).

```{r santa_clara_county, fig.height=4, fig.width=8, echo = FALSE}
scc_2020 = adp_girls %>%
  filter(county == "Santa Clara" & year == 2020) %>%
  mutate(
    month = as.factor(month),
    day = as.numeric(day),
    year = as.numeric(year),
    ADP = total_female)

plot1 = scc_2020 %>%
  ggplot(aes(x = month, y = ADP, fill = ADP)) +
  geom_bar(position="dodge", stat="identity") +
    theme(legend.position=c(0.9, 0.8)) +
  labs(
    title = "Santa Clara County ADP, 2020",
    x = "Month",
    y = "ADP") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))


scc_2021 = adp_girls %>%
  filter(county == "Santa Clara" & year == 2021) %>%
  mutate(
    month = as.factor(month),
    day = as.numeric(day),
    year = as.numeric(year),
    ADP = total_female)

plot2 = scc_2021 %>%
  ggplot(aes(x = month, y = ADP, fill = ADP)) +
  geom_bar(position="dodge", stat="identity") +
    theme(legend.position=c(0.9, 0.8)) +
  labs(
    title = "Santa Clara County ADP, 2021",
    x = "Month") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))

plot1 + plot2

```

# DETENTION PROFILES

The California Juvenile Court and Probation Statistical System (JCPSS) reports data on the proportion of cases within racial groups that result in types of detention decisions.

<br>

```{r detentions, echo = FALSE, message = FALSE, warning = FALSE, fig.height=10, fig.width=9}
detention_df = read_csv("/Users/shainamackin/Desktop/Vera/California Criminalization/data/statewide detention by race.csv") %>%
  select(prop.col.y, prop.col.Freq, prop.tbl.x) %>%
  drop_na() %>%
  janitor::clean_names() %>%
  mutate(
    race = prop_col_y,
    detainment = prop_tbl_x,
    proportion = prop_col_freq) %>%
  mutate(
    detainment = replace(detainment, detainment == 'Detained - Home Supervision', 'Home Supervision'),
    detainment = replace(detainment, detainment == 'Detained - Non-Secure Facility', 'Non-Secure Facility'),
      detainment = replace(detainment, detainment == 'Detained - Secure Facility', 'Secure Facility')) %>%
  select(-prop_col_y, -prop_tbl_x, -prop_col_freq)

ggplot(detention_df, aes(x = reorder(detainment, proportion), y = proportion, fill = detainment)) +
  geom_bar(stat='identity',position="dodge") +
  coord_flip() +
  facet_wrap(~race, ncol=4) + 
    labs(
    title = "Statewide Detention Profiles by Race, 2020",
    x = "Detainment Decision",
    y = "Proportion") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label=sprintf("%0.0f", round(proportion))))
```


```{r detentions2, include = FALSE, message = FALSE, warning = FALSE, fig.height=8, fig.width=9}
ggplot(detention_df, aes(x = reorder(race, -proportion), y = proportion, fill = race)) +
  geom_bar(stat='identity',position="dodge") +
  facet_wrap(~detainment, ncol=1) + 
    labs(
    title = "Statewide Detention Profiles by Race, 2020",
    x = "Detainment Decision",
    y = "Proportion") +
  theme(axis.text.x=element_text(angle = 90, hjust = 1))
```
<br>


<br>
<br>
<br>

# ARREST RATES

```{r arrests, echo = FALSE, message = FALSE, warning = FALSE}
```

